The Hyperedge Event Model

### 2018-06-27 paper/ba-HEM/ba-sample.tex
rearranged figures

	\subsection{Senders, receivers, and timestamps}\label{subsec:Observed}
	Finally, our model assumes that the observed sender, receivers, and timestamp of hyperedge event $e$ are generated by selecting the sender--receiver-set pair with the smallest time increment \citep{snijders1996stochastic}:
	\begin{equation}
		\begin{aligned}
			s_e &= \mbox{argmin}_{i}(\tau_{ie}),\\
			\boldsymbol{r}_e &= \boldsymbol{u}_{s_e e},\\
			t_e &=t_{e-1} + \tau_{s_e e}.
		\end{aligned}
	\end{equation}
	\begin{figure}[!t]
		\centering
		\includegraphics[width=0.625\textwidth]{img/diagramnew.png}	
		\caption {An illustrative example of the generative process of the HEM.}
		\label{figure:diagram}
	\end{figure}	
		Therefore, the HEM assumes a sender-driven process---i.e., the receivers and timestamp of an event are jointly determined by the sender's urgency to direct the event to those selected receivers. Note that our generative process allows for tied events. In the case of tied events---i.e., multiple senders draw exactly the same candidate timestamps---we assume that all events are generated and occur simultaneously. Algorithm \ref{alg:generative} summarizes the entire generative process for hyperedge events with one sender and one or more receivers, and Figure \ref{figure:diagram} presents an illustrative example on how the $e^{\textrm{th}}$ event is generated, assuming $t_{e-1} = 0$ and $A=5$.~\looseness=-1
	\begin{algorithm}[!t]
		\spacingset{1}
		\SetAlgoLined
		\caption{Generative process: one sender and one or more receivers}
		\begin{algorithmic}
			\STATE \textbf{Input}: number of events and nodes $(E, A)$, covariates $(\boldsymbol{x}, \boldsymbol{y})$, and coefficients $(\boldsymbol{b}, \boldsymbol{\eta})$
			\vskip 0.1in
			\FOR{$e=1$ to $E$}
			\FOR{$i=1$ to $A$}
			\FOR{$j=1$ to $A$ ($j \neq i$)}
			\STATE	set $\lambda_{iej} = {\boldsymbol{b}}^{\top}\boldsymbol{x}_{iej}$
			\ENDFOR
			\STATE	draw $\boldsymbol{u}_{ie}  \sim
			\mbox{MB}_G(\boldsymbol{\lambda}_{ie})$
			\STATE		set $\mu_{ie} = g^{-1}(\boldsymbol{\eta}^\top \boldsymbol{y}_{ie})$
			\STATE		draw $\tau_{ie} \sim f_\tau(\mu_{ie}, V(\mu))$
			\ENDFOR

			\IF{$n \geq 2$ tied events} 
			\STATE	set $s_e,\ldots, s_{e+n-1}=\mbox{argmin}_{i}(\tau_{ie}),$
			\STATE	set $\boldsymbol{r}_e=\boldsymbol{u}_{s_e e},\ldots,\boldsymbol{r}_{e+n-1}=\boldsymbol{u}_{s_{e+n-1} e}$
			\STATE	set $t_e, \ldots, t_{e+n-1}=t_{e-1} + \min_i\tau_{ie}$
			\STATE		jump to $e = e+n$
			\ELSE
			\STATE	set $s_e= \mbox{argmin}_{i}(\tau_{ie})$
			\STATE		set $\boldsymbol{r}_e = \boldsymbol{u}_{s_e e}$
			\STATE	set $t_e =t_{e-1} + \min_i\tau_{ie}$
			\ENDIF
			\ENDFOR
		\end{algorithmic}
		\label{alg:generative}
	\end{algorithm}

@@

	\subsection{Getting it Right (GiR) test} \label{subec:GiR}
	Software development is integral to the objective of applying our model to real world data. Code review is a valuable process in any research computing context, and the prevalence of software bugs in statistical software is well documented \citep[e.g., ][]{altman2004numerical,mccullough2009accuracy}.  With highly complex models such as the HEM, there are many ways in which software bugs can be introduced and go unnoticed. As such, we present a joint analysis of the integrity of our generative model, sampling equations, and software implementation.

@@

	In the forward sampler, both observable and unobservable variables are generated using Algorithm \ref{alg:generative}. In the backward samples, unobservable variables are generated using the sampling equations for inference, which we derived in Section \ref{subsec:conditionaldist}. For each forward and backward sample that consists of $E$ number of events, we save these statistics:
	\begin{itemize}
		\item[1.] Mean of observed receiver sizes $ \lVert \boldsymbol{r}_{e} \rVert_1 $ across $e=1,\ldots,E$,
		\item[2.] Variance of observed receiver sizes $ \lVert \boldsymbol{r}_{e} \rVert_1 $ across $e=1,\ldots,E$,
		\item[3.] Mean of time increments $\tau_e$ across $e=1,...,E$,
		\item[4.] Variance of time increments $\tau_e$ across $e=1,...,E$,
		\item[5.] $b_p$ value used to generate the samples $p = 1,...,P$,
		\item[6.] $\eta_q$ value used to generate the samples $q = 1,...,Q$,
		\item[7.] $\sigma^2_\tau$ value used to generate the samples in log-normal distribution
	\end{itemize}

@@

	\subsubsection{Coefficients for receiver selection features}
	Figure \ref{figure:betaresults} shows the boxplots summarizing posterior samples of $\boldsymbol{b}$, where Figure \ref{subfigure:betaresults1} displays the coefficients for nodal covariates and \ref{subfigure:betaresults2} displays the coefficients for dyadic and triadic covariates. Since we use the logit functional form 
	\begin{equation*}
		\mbox{logit}(\lambda_{iej})=\log\Big(\frac{\lambda_{iej}}{1-\lambda_{iej}}\Big) =b_{1}+b_{2} x_{iej2}\ldots+b_{14}x_{iej14},
	\end{equation*}
	and can interpret the $\boldsymbol{b}$ estimates in terms of odds ratios $\frac{\lambda_{iej}}{1-\lambda_{iej}}=\exp(b_{1}+b_{2} x_{iej2}\ldots+b_{14}x_{iej14})$. First of all, we find the effects of nodal coavariates ``gender\_\,sender$(i)$" and ``gender\_\,receiver$(j)$" are both nearly always negative in the posterior samples. The log odds that any other node will be added as a receiver of an email is approximately two times less if the sender is a woman. The posterior distribution of the statistic ``outdegree$(i)$" is mostly negative, if sender $i$ sent $n$ number of emails to anyone last week, then sender $i$ is approximately $\exp(-0.109\times n)\approx(0.897)^n$ times less likely to send an email to $j$. However, this straightforward interpretation of the outdegree statistic only applies when the hyperedge size is low. The scenario in which a sender sends a lot of low-hyperedge-size emails may arise due to the use of email for a one-on-one conversation. The large positive estimates of the interaction between hyperedge size and outdegree indicate that those who have recently sent many emails with many receivers on each email are likely to continue sending broadcast emails. This scenario may arise from someone being responsible for distributing timely announcements. When we look at the effect of ``indegree$(j)$," we see a clear popularity effect---those who have received a lot of emails a lot recently are likely to continue receiving a lot of emails. If the receiver $j$ received $n$ number of emails over the last week, sender $i$ is $\exp(0.086\times n)\approx(1.091)^n $ times more likely to send an email to $j$.

@@

				\begin{tabular}[t]{cc}
					\begin{subfigure}[b]{0.4975\textwidth}
						\caption{Nodal covariates}
						\includegraphics[width=\textwidth]{img/betanewplot2-1.png}
						\label{subfigure:betaresults1}	
					\begin{subfigure}[b]{0.4975\textwidth}
						\caption{Dyadic and triadic covariates}
						\includegraphics[width=\textwidth]{img/betanewplot3-1.png}
						\label{subfigure:betaresults2}	
				\caption {Posterior distribution of $\boldsymbol{b}$ estimates.}
				\label{figure:betaresults}

@@

		\iffalse
		\begin{supplement}
			\sname{Supplement A}\label{suppA} 
			\stitle{Title of the Supplement A}
			\slink[url]{http://www.some-url-address.org/dowload/0000.zip}
			\sdescription{Add description for supplement material.}
		\end{supplement}
		\fi

		\bibliographystyle{ba}
		\bibliography{baHEM}

>>>>

	\subsection{Senders, receivers, and timestamps}\label{subsec:Observed}
	Finally, our model assumes that the observed sender, receivers, and timestamp of hyperedge event $e$ are generated by selecting the sender--receiver-set pair with the smallest time increment \citep{snijders1996stochastic}:
	\begin{equation}
		\begin{aligned}
			s_e &= \mbox{argmin}_{i}(\tau_{ie}),\\
			\boldsymbol{r}_e &= \boldsymbol{u}_{s_e e},\\
			t_e &=t_{e-1} + \tau_{s_e e}.
		\end{aligned}
	\end{equation}
	\begin{algorithm}[!t]
		\spacingset{1}
		\SetAlgoLined
		\caption{Generative process: one sender and one or more receivers}
		\begin{algorithmic}
			\STATE \textbf{Input}: number of events and nodes $(E, A)$, covariates $(\boldsymbol{x}, \boldsymbol{y})$, and coefficients $(\boldsymbol{b}, \boldsymbol{\eta})$
			\vskip 0.1in
			\FOR{$e=1$ to $E$}
			\FOR{$i=1$ to $A$}
			\FOR{$j=1$ to $A$ ($j \neq i$)}
			\STATE	set $\lambda_{iej} = {\boldsymbol{b}}^{\top}\boldsymbol{x}_{iej}$
			\ENDFOR
			\STATE	draw $\boldsymbol{u}_{ie}  \sim
			\mbox{MB}_G(\boldsymbol{\lambda}_{ie})$
			\STATE		set $\mu_{ie} = g^{-1}(\boldsymbol{\eta}^\top \boldsymbol{y}_{ie})$
			\STATE		draw $\tau_{ie} \sim f_\tau(\mu_{ie}, V(\mu))$
			\ENDFOR

			\IF{$n \geq 2$ tied events} 
			\STATE	set $s_e,\ldots, s_{e+n-1}=\mbox{argmin}_{i}(\tau_{ie}),$
			\STATE	set $\boldsymbol{r}_e=\boldsymbol{u}_{s_e e},\ldots,\boldsymbol{r}_{e+n-1}=\boldsymbol{u}_{s_{e+n-1} e}$
			\STATE	set $t_e, \ldots, t_{e+n-1}=t_{e-1} + \min_i\tau_{ie}$
			\STATE		jump to $e = e+n$
			\ELSE
			\STATE	set $s_e= \mbox{argmin}_{i}(\tau_{ie})$
			\STATE		set $\boldsymbol{r}_e = \boldsymbol{u}_{s_e e}$
			\STATE	set $t_e =t_{e-1} + \min_i\tau_{ie}$
			\ENDIF
			\ENDFOR
		\end{algorithmic}
		\label{alg:generative}
	\end{algorithm}
		Therefore, the HEM assumes a sender-driven process---i.e., the receivers and timestamp of an event are jointly determined by the sender's urgency to direct the event to those selected receivers. Note that our generative process allows for tied events. In the case of tied events---i.e., multiple senders draw exactly the same candidate timestamps---we assume that all events are generated and occur simultaneously. Algorithm \ref{alg:generative} summarizes the entire generative process for hyperedge events with one sender and one or more receivers, and Figure \ref{figure:diagram} presents an illustrative example on how the $e^{\textrm{th}}$ event is generated, assuming $t_{e-1} = 0$ and $A=5$.~\looseness=-1	
		\begin{figure}[!t]
			\centering
			\includegraphics[width=0.625\textwidth]{img/diagramnew.png}	
			\caption {An illustrative example of the generative process of the HEM.}
			\label{figure:diagram}
		\end{figure}

@@

	\subsection{Getting it Right (GiR) test} \label{subec:GiR}
	Software development is integral to the objective of applying our model to real world data. Code review is a valuable process in any research computing context, and the prevalence of software bugs in statistical software is well documented \citep[e.g., ][]{altman2004numerical,mccullough2009accuracy}.  With highly complex models such as the HEM, there are many ways in which software bugs can be introduced and go unnoticed. As such, we present a joint analysis of the integrity of our generative model, sampling equations, and software implementation.

@@

	In the forward sampler, both observable and unobservable variables are generated using Algorithm \ref{alg:generative}. In the backward samples, unobservable variables are generated using the sampling equations for inference, which we derived in Section \ref{subsec:conditionaldist}. For each forward and backward sample that consists of $E$ number of events, we save these statistics:\\
	\begin{itemize}
		\item[1.] Mean of observed receiver sizes $ \lVert \boldsymbol{r}_{e} \rVert_1 $ across $e=1,\ldots,E$,
		\item[2.] Variance of observed receiver sizes $ \lVert \boldsymbol{r}_{e} \rVert_1 $ across $e=1,\ldots,E$,
		\item[3.] Mean of time increments $\tau_e$ across $e=1,...,E$,
		\item[4.] Variance of time increments $\tau_e$ across $e=1,...,E$,
		\item[5.] $b_p$ value used to generate the samples $p = 1,...,P$,
		\item[6.] $\eta_q$ value used to generate the samples $q = 1,...,Q$,
		\item[7.] $\sigma^2_\tau$ value used to generate the samples in log-normal distribution
	\end{itemize}

@@

	\subsubsection{Coefficients for receiver selection features}
	Figure \ref{figure:betaresults} shows the boxplots summarizing posterior samples of $\boldsymbol{b}$, where Figure \ref{subfigure:betaresults1} displays the coefficients for nodal covariates and \ref{subfigure:betaresults2} displays the coefficients for dyadic and triadic covariates. Since we use the logit functional form 
	\begin{equation*}
		\mbox{logit}(\lambda_{iej})=\log\Big(\frac{\lambda_{iej}}{1-\lambda_{iej}}\Big) =b_{1}+b_{2} x_{iej2}\ldots+b_{14}x_{iej14},
	\end{equation*}
	and can interpret the $\boldsymbol{b}$ estimates in terms of odds ratios $\frac{\lambda_{iej}}{1-\lambda_{iej}}=\exp(b_{1}+b_{2} x_{iej2}\ldots+b_{14}x_{iej14})$. First of all, we find the effects of nodal coavariates ``gender\_\,sender$(i)$" and ``gender\_\,receiver$(j)$" are both nearly always negative in the posterior samples. The log odds that any other node will be added as a receiver of an email is approximately two times less if the sender is a woman. The posterior distribution of the statistic ``outdegree$(i)$" is mostly negative, if sender $i$ sent $n$ number of emails to anyone last week, then sender $i$ is approximately $\exp(-0.109\times n)\approx(0.897)^n$ times less likely to send an email to $j$. However, this straightforward interpretation of the outdegree statistic only applies when the hyperedge size is low. The scenario in which a sender sends a lot of low-hyperedge-size emails may arise due to the use of email for a one-on-one conversation. The large positive estimates of the interaction between hyperedge size and outdegree indicate that those who have recently sent many emails with many receivers on each email are likely to continue sending broadcast emails. This scenario may arise from someone being responsible for distributing timely announcements. When we look at the effect of ``indegree$(j)$," we see a clear popularity effect---those who have received a lot of emails a lot recently are likely to continue receiving a lot of emails. If the receiver $j$ received $n$ number of emails over the last week, sender $i$ is $\exp(0.086\times n)\approx(1.091)^n $ times more likely to send an email to $j$. 
			\begin{figure}[!t]
				\centering
				\begin{tabular}[t]{cc}
					\begin{subfigure}[b]{0.4975\textwidth}
						\caption{Nodal covariates}
						\includegraphics[width=\textwidth]{img/betanewplot2-1.png}
						\label{subfigure:betaresults1}	
					\end{subfigure}
					\begin{subfigure}[b]{0.4975\textwidth}
						\caption{Dyadic and triadic covariates}
						\includegraphics[width=\textwidth]{img/betanewplot3-1.png}
						\label{subfigure:betaresults2}	
					\end{subfigure}
				\end{tabular}
				\caption {Posterior distribution of $\boldsymbol{b}$ estimates.}
				\label{figure:betaresults}
			\end{figure}

@@

@@

		\iffalse
		\begin{supplement}
			\sname{Supplement A}\label{suppA} 
			\stitle{Title of the Supplement A}
			\slink[url]{http://www.some-url-address.org/dowload/0000.zip}
			\sdescription{Add description for supplement material.}
		\end{supplement}
		\fi
		\newpage
		\bibliographystyle{ba}
		\bibliography{baHEM}
